A Kochen-Specker system has at least 22 vectors

### 2015-05-21 paper/main.tex
paper: second cross product unemb. proof

\newcommand{\keyword}[1]{\textbf{#1}}

@@

            ;
        \end{tikzpicture}
    \end{center}
    These three graphs are unembeddable.  The left and middle graph
    are the only minimal unembeddable squarefree graph.
For the first graph, we have proven directly that it is unembeddable.
See Proposition~\ref{prop:unemb-10-2}.
For the second graph, we also have a similar direct proof. The third graph is shown to not be embeddable using our algorithm.
\end{comp}

@@

\appendix
\clearpage
\section{Appendix}
In this appendix, we show by hand, that indeed the 19 vertex graph
considered at the start of section 3 %TODO
is not embeddable.
We restrict the calculation to the~$13$-vertex unembeddable
subgraph.

@@

>>>>

\newcommand{\keyword}[1]{\textbf{#1}}
\newcommand{\coll}{\mathrel{\text{coll.}}}

@@

            ;
        \end{tikzpicture}
    \end{center}
    These three graphs are unembeddable.  The left and middle graph
    are the only minimal unembeddable squarefree graph.
For the first graph, we have proven directly that it is unembeddable.
See Proposition~\ref{prop:unemb-10-2}.
For the second graph, we also have a similar direct proof.
See Proposition~\ref{prop:unemb-10-1} in the appendix.
The third graph is shown to not be embeddable using our
algorithm.
\end{comp}

@@

% Apparently, \appendix does not work properly.
% Thus, we set \thesubsection by hand
\appendix
\clearpage
\renewcommand{\thesubsection}{\Alph{subsection}}
\renewcommand{\thesection}{\Alph{subsection}}
\section*{Appendix}
\subsection{Original pen-and-paper proof}
In this appendix, we show by hand, that indeed the 19 vertex graph
considered at the start of section 3 %TODO
is not embeddable.
We restrict the calculation to the~$13$-vertex unembeddable
subgraph.

@@

\subsection{Second cross-product unembeddability proof}

\parpic(5cm,3.5cm)[r]{%
\begin{tikzpicture}[thick,scale=1.4,
    p/.style={circle, draw, fill=white,
                    inner sep=0pt, minimum width=5pt}]
        \draw
            \foreach \x in {0,180} {
                (\x+15:1.3) node[p]{} -- (\x+90:0.9) node[p]{}
                (\x+15:1.3) -- (\x+45:0.5) node[p]{}
                (\x+45:0.5) -- (\x+90:0.9)
                (\x-15:1.3) node[p]{} -- (\x-90:0.9)
                (\x-15:1.3) -- (\x-45:0.5) node[p]{}
                (\x-45:0.5) -- (\x-90:0.9)
                (\x-15:1.3) -- (\x+15:1.3)
            }
            (45:0.5) -- (45+180:0.5)
            (45+90:0.5) -- (45+90+180:0.5)
            (90:0.9) node[above] {$x$}
            (270:0.9) node[below] {$w$}
            (15:1.3) node[right] {$y$}
            (-15:1.3) node[right] {$v$}
            (45:0.5) node[below] {$z$}
            (-45:0.5) node[above] {$p_5$}
            (135:0.5) node[below] {$p_1$}
            (-135:0.5) node[above] {$p_4$}
            (165:1.3) node[left] {$p_2$}
            (-165:1.3) node[left] {$p_3$}
        ;
    \end{tikzpicture}}

\par\noindent\parbox{\textwidth}{%
\begin{proposition}\label{prop:unemb-10-1}
The graph on the right is not embeddable.
\end{proposition}}
\begin{myproof}
    Suppose it is embeddable.
    Note (the point associated to) $p_5$
        is orthogonal to (the points associated to) $v$ and~$w$.
    As~$v \neq w$, we see~$p_5$ is collinear
    to cross-product~$v \times w$.
    Similarly, we derive
    \begin{align*}
        p_1 &\coll x \times (v \times w) &
        p_4 &\coll w \times z \\
        p_2 &\coll x \times (x \times v \times w) &
        p_3 & \coll w \times (w \times z) \\
        \omit\rlap{$
        p_3 \coll (w \times z) \times (x \times (x \times (v \times w)))$}
    \end{align*}
    And thus
    \begin{equation}
        (w \times (w \times z)) \times
        ((w \times z) \times (x \times (x\times (v \times w)))) =0.
            \label{eq:coll2}
    \end{equation}
    By rotating the embedding,
        we may assume without loss of generality
            that we have~$z=(0,0,1)$, $y=(0,1,0)$ and~$x=(1,0,0)$.
    Write
\begin{align*}
    v&=\Bigl(v_1,0,\sqrt{1-v_1^2}\Bigr) &
        w &= \Bigl(w_1,w_2,\sqrt{1-w_1^2-w_2^2}\Bigr).
\end{align*}
Now, equation~\ref{eq:coll2} becomes
\begin{align*}
    w_2^2 \Bigl(v_1-\sqrt{1-v_1^2}w_1\sqrt{1-w_1^2-w_2^2}\Bigr) &= 0 \\
    w_1w_2\Bigl(\sqrt{1-v_1^2}w_1 \sqrt{1-w_1^2 - w_2^2} - v_1 \Bigr) &= 0.
\end{align*}
Note~$w_1 \neq 0$,
for otherwise~$w$ would be orthogonal to~$x$,
which would imply~$w = y$, quod non.
Similarly~$w_2 \neq 0$, for otherwise~$w$ would be orthogonal to~$y$
and then~$v = p_4$, quod non.
And finally, $v_1 \neq 0$, for otherwise~$v$ would be orthogonal to~$x$
and then~$x = p_5$, quod non.
It follows~$v_1 - \sqrt{1-v_1^2}  w_1 \sqrt{1-w_1^2 - w_2^2} = 0$.
Thus $v_1 = \sqrt{1-v_1^2}  w_1 \sqrt{1-w_1^2 - w_2^2}$.
As $v$ is orthogonal to~$w$,
we have~$-v_1 w_1 = \sqrt{1-v_1^2}\sqrt{1-w_1^2-w_2^2}$.
Substituting this in the previous equation,
we get~$v_1 = -v_1w_1^2$.
That is: $w_1^2 = -1$, which is absurd.
Apparently, the graph cannot be embeddable.
\end{myproof}
